An equivalence of compositional frameworks

### 2018-10-03 EquivCompFrameworks_20180911kc.tex
Changed notation to match notation used in structured cospans paper

\begin{abstract}
\noindent
The first two authors have developed a compositional framework well-suited for studying networks that are built out of finite sets equipped with extra stuff. This framework, which goes by the name of `structured cospans', utilizes double categories where the objects are seen as inputs and outputs, morphisms are `open networks', and 2-morphisms are maps between open networks. In this setup a functor $L \colon \bold{C} \to \bold{D}$, which is typically a left adjoint, is used to replace the objects and vertical 1-morphisms of a given double category $\mathbb{D}$ with the objects and morphisms, respectively, of the category $\bold{C}$. Horizontal 1-cells are then cospans in $\bold{D}$ of a particular form with 2-morphisms given by maps of these cospans. Fong has also developed a similar framework which also uses cospans and goes by the name of `decorated cospans'. In this setup, a lax monoidal functor $F \colon \bold{C} \to \bold{E}$ is used to `decorate' the apices of cospans in $\bold{C}$ with elements of $\bold{E}$ giving the cospans extra structure. Using a slight variation of Fong's framework, we prove that these two frameworks are equivalent in the situation where a left adjoint can be obtained from a lax monoidal pseudofunctor using a well known construction of Grothendieck.
\end{abstract}

\section{Introduction}

git pro coming through!

Networks are playing an increasingly prominent role in our understanding of the world and as a result, methods of characterizing and studying networks have become evermore necessary. Applied category theory provides such an avenue and much work has been done towards this effort  \cite{BC,BCR,BF,BFP,BP,Cour,Fong,JM,LS,Yass}. One of the more recent frameworks developed suitable for studying networks is Fong's `decorated cospans' \cite{Fong}. A $\textbf{cospan}$ in any category is a diagram of the form:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,1) {$c$};
\node (C) at (2,0) {$b$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
We call $c$ the $\textbf{apex}$ of the cospan and $a$ and $b$ the $\textbf{feet}$ of the cospan. The morphisms $i$ and $o$ are called the $\textbf{legs}$. Cospans are ideal for realizing networks and more generally `open' systems where here open means each system or network comes with prescribed inputs and outputs which are represented by the feet of the cospan $a$ and $b$, respectively. Two open systems viewed as cospans such that the outputs of the first and the inputs of the second coincide can then naturally be composed via pushout. For example, we can compose the above cospan with the following:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$b$};
\node (B) at (1,1) {$f$};
\node (C) at (2,0) {$e$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i^\prime$} (B)
(C) edge node[above]{$o^\prime$} (B);
\end{tikzpicture}
\]
to obtain a new cospan whose left foot is the left foot of the first, whose right foot is the right foot of the second, and whose apex and legs are given by the pushout.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,1) {$c$};
\node (C) at (2,0) {$b$};
\node (D) at (3,1) {$f$};
\node (E) at (4,0) {$e$};
\node (F) at (2,2) {$c+_b f$};
\node (G) at (2,1.25) {$c+f$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(C) edge node[above]{$i^\prime$} (D)
(E) edge node[above]{$o^\prime$} (D)
(B) edge node[left]{$\psi j_c$} (F)
(D) edge node[right]{$\psi j_f$} (F)
(B) edge node[above]{$j_c$} (G)
(D) edge node[above]{$j_f$} (G)
(G) edge node [left] {$\psi$} (F);
\end{tikzpicture}
\]

In Fong's theory of decorated cospans, given a category $\bold{C}$ with finite colmits, a symmetric lax monoidal functor $F \colon (\bold{C},+,0) \to (\bold{D},\otimes,I)$ is used to `decorate' the apex of a cospan in $\bold{C}$ with an element of the image of its apex under the functor $F$, in the case of the first cospan above, a morphism $d \colon I \to F(c)$. Here $F(c)$ is the collection of all $F$-decorations on the object $c$ and the morphism $d$ is selecting a particular one. Thus a decorated cospan is given by a pair:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c$};
\node (C) at (2,0) {$b$};
\node (D) at (3,0) {$I$};
\node (E) at (4,0) {$F(c)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(D) edge node [above] {$d$} (E);
\end{tikzpicture}
\]
To obtain a decoration on the composition of two composable $F$-decorated cospans:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c$};
\node (C) at (2,0) {$b$};
\node (F) at (3,0) {$b$};
\node (G) at (4,0) {$f$};
\node (H) at (5,0) {$e$};
\node (D) at (0.5,-0.5) {$I$};
\node (E) at (1.5,-0.5) {$F(c)$};
\node (D') at (3.5,-0.5) {$I$};
\node (E') at (4.5,-0.5) {$F(f)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(F) edge node[above]{$i^\prime$} (G)
(H) edge node[above]{$o^\prime$} (G)
(D) edge node [above] {$d_1$} (E)
(D') edge node [above] {$d_2$} (E');
\end{tikzpicture}
\]
we use the natural map from a coproduct to a pushout as well as the natural map that comes as part of the structure of a lax monoidal functor:
 $$1 \xrightarrow{\lambda^{-1}} 1 \times 1 \xrightarrow{d_1 \times d_2} F(c) \times F(f) \xrightarrow{\phi_{c,f}} F(c+ f) \xrightarrow{F(j_{c,f})} F(c+_b f).$$

From this symmetric lax monoidal functor $F \colon \bold{C} \to \bold{D}$, Fong then constructs a symmetric monoidal category $F\textnormal{Cospan}(\bold{C})$ which has:
\begin{enumerate}
\item{objects as those of $\bold{C}$ and}
\item{morphisms as isomorphism classes of $F$-decorated cospans, where an $F$-decorated cospan is given as above, and two $F$-decorated cospans are in the same isomorphism class if there exists an isomorphism between the apices such that the following diagrams commute:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,1) {$c$};
\node (B') at (1,-1) {$c^\prime$};
\node (C) at (2,0) {$b$};
\node (D) at (3,0) {$I$};
\node (E) at (4,0.5) {$F(c)$};
\node (E') at (4,-0.5) {$F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[below]{$i^\prime$} (B')
(C) edge node[below]{$o^\prime$} (B')
(B) edge node [left] {$f$} (B')
(D) edge node [above] {$d$} (E)
(D) edge node [below] {$d^\prime$} (E')
(E) edge node [right] {$F(f)$} (E');
\end{tikzpicture}
\]
}
\end{enumerate}
There are some subtleties to this approach. To illustrate, consider the example where $F \colon \bold{FinSet} \to \bold{Set}$ is the symmetric lax monoidal functor that assigns to a finite set $N$ the (large) set of all possible graph structures on the finite set $N$, where a graph structure on $N$ is given by a diagram in $\bold{Set}$ of the form:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$E$};
\node (B) at (1,0) {$N.$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge[bend left] node[above]{$s$} (B)
(A) edge[bend right] node[below]{$t$} (B);
\end{tikzpicture}
\]
Let $N=\{ v_1,v_2 \}$ be a two element set. Then one element of the set $F(N)$, in other words, one possible decoration, which is a graph structure, on the set $N$ is given by a single edge $e$ whose source and target are $v_1$ and $v_2$, respectively.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$v_1$};
\node (B) at (1,0) {$v_2$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$e$} (B);
\end{tikzpicture}
\]
Denote this element of $F(N)$ as $d$. We could label the edge $e$ something else, for instance, $e^\prime$, and this single edge would another element of $F(N)$.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$v_1$};
\node (B) at (1,0) {$v_2$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$e^\prime$} (B);
\end{tikzpicture}
\]
Denote this second element of $F(N)$ as $d^\prime$. These two graphs constitute distinct isomorphism classes in the symmetric monoidal category $F$Cospan$(\bold{FinSet})$, as $f=\id_N$ and thus $F(f)=\id_{F(N)}$, but clearly $d \neq d^\prime$. The issue here is that the functor $F$ is decorating each finite set $N$ not only with extra structure, but with extra \emph{stuff}. Moreover, the finite set $N$ is being decorated with \emph{different} stuff in each instance and this difference in stuff is not entirely detected by the function $F(f)$. This phenomenon doesn't occur when the decorations only involve extra structure, and this has been utilized in other works \cite{BFP,BP,Yass}.

One approach to remedying this nuisance is to instead view $F$ as a functor $F \colon\bold{C} \to \bold{Cat}$ and then consider $F(c)$ not as a \emph{set} of decorations on the element $c$ but as a \emph{category} of decorations on the element $c$. Then one can construct a similar symmetric monoidal category similar to Fong's, which by similarity and an abuse of notation we also call $F\textnormal{Cospan}(\bold{C})$. This symmetric monoidal category will have:
\begin{enumerate}
\item{objects as those of $\bold{C}$ and}
\item{morphisms as isomorphism classes of cospans of $\bold{C}$ together with an element of image of the apex under the functor $F$.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c$};
\node (C) at (2,0) {$b$};
\node (D) at (3,0) {$d \in F(c)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
Except now, given another morphism:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c^\prime$};
\node (C) at (2,0) {$b$};
\node (D) at (3,0) {$d^\prime \in F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i^\prime$} (B)
(C) edge node[above]{$o^\prime$} (B);
\end{tikzpicture}
\]
these two morphisms are in the same isomorphism class if the following diagrams commute
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,1) {$c$};
\node (B') at (1,-1) {$c^\prime$};
\node (C) at (2,0) {$b$};
\node (D) at (3,0) {$\bold{1}$};
\node (E) at (4,0.5) {$F(c)$};
\node (E') at (4,-0.5) {$F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[below]{$i^\prime$} (B')
(C) edge node[below]{$o^\prime$} (B')
(B) edge node [left] {$f$} (B')
(D) edge node [above] {$d$} (E)
(D) edge node [below] {$d^\prime$} (E')
(E) edge node [right] {$F(f)$} (E');
\end{tikzpicture}
\]
and there exists an isomorphism $\iota \colon F(f)(d) \to d^\prime$ in the category $F(c^\prime)$.
}
\end{enumerate}
Note that the existence of an isomorphism $\iota \colon F(f)(d) \to d^\prime$ is equivalent to the triangle above on the right commuting up to a 2-isomorphism.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (4,0) {$\iota \Swarrow$};
\node (D) at (3,0) {$\bold{1}$};
\node (E) at (4.5,1) {$F(c)$};
\node (E') at (4.5,-1) {$F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(D) edge node [above] {$d$} (E)
(D) edge node [below] {$d^\prime$} (E')
(E) edge node [right] {$F(f)$} (E');
\end{tikzpicture}
\]
In other words, the isomorphism $\iota$ is a \emph{natural} isomorphism. We will typically just write this as $\iota \colon F(f)(d) \to d^\prime$ to conserve space.

In this new category, the added structure of an isomorphism in the category of decorations shapes the isomorphism classes to look as one would expect. Fong's symmetric monoidal category has been extended to a bicategory by the first author \cite{Cour} and this was one of the original illuminations of the above obstacle. In this bicategory, objects and morphisms are given as in Fong's category modulo taking cospans up to isomorphism class and 2-morphisms are given by a pair of commuting diagrams:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,1) {$c$};
\node (B') at (1,-1) {$c^\prime$};
\node (C) at (2,0) {$b$};
\node (D) at (3,0) {$\bold{1}$};
\node (E) at (4,0.5) {$F(c)$};
\node (E') at (4,-0.5) {$F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[below]{$i^\prime$} (B')
(C) edge node[below]{$o^\prime$} (B')
(B) edge node [left] {$f$} (B')
(D) edge node [above] {$d$} (E)
(D) edge node [below] {$d^\prime$} (E')
(E) edge node [right] {$F(f)$} (E');
\end{tikzpicture}
\]
In this case, there would be no 2-morphism from:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$v_1$};
\node (B) at (1,0) {$v_2$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$e$} (B);
\end{tikzpicture}
\]
to:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$v_1$};
\node (B) at (1,0) {$v_2.$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$e^\prime$} (B);
\end{tikzpicture}
\]
Morally speaking, these two simple single-edged graphs should belong to the same isomorphism class, as there is a natural isomorphism of graphs between them.

In the present work, we construct a symmetric monoidal `double category' $\mathbb{F}\textnormal{Cospan}(\bold{C})$. Double categories were first introduced by Ehresmann \cite{Ehresmann63, Ehresmann65} and symmetric monoidal double categories by Shulman \cite{Shul}. They have long been used in topology and other branches of pure mathematics \cite{Brown1,Brown2}.  More recently they have been used to study open dynamical systems \cite{LS} and open Markov processes \cite{BC}. While a mere category has only objects and morphisms, a double category has a few more types of entities:

@@

We call $A, B, C$ and $D$ `objects', $f$ and $g$ `vertical 1-morphisms', $M$ and $N$ `horizontal 1-cells', and $a$ a `2-morphism'.   We can compose vertical 1-morphisms to get new vertical 1-morphisms and compose horizontal 1-cells to get new horizontal 1-cells.  We can compose the 2-morphisms in two ways: horizontally by setting squares side by side, and vertically by setting one on top of the other.   In a `strict' double category all these forms of composition are associative.  In a `pseudo' double category, horizontal 1-cells compose in a weakly associative manner: that is, the associative law holds only up to an invertible 2-morphism, called the `associator', which obeys a coherence law. The double category $\mathbb{F}\textnormal{Cospan}(\bold{C})$ that we construct in this paper has:

@@

\item{objects as those of $\bold{C}$,}
\item{vertical 1-morphisms as morphisms of $\bold{C}$,}
\item{horizontal 1-cells as $F$-decorated cospans of $\bold{C}$, which are pairs:

@@

We can then obtain a symmetric monoidal bicategory $H(\mathbb{F}\textnormal{Cospan}(\bold{C}))$ as the `horizontal bicategory' of the symmetric monoidal double category $\mathbb{F}\textnormal{Cospan}(\bold{C})$ using a result of Shulman \cite{Shul}. This bicategory will have:

@@

\item{objects as those of $\bold{C}$,}

@@

\node (D) at (3,0) {$\bold{1}$};

@@

\end{enumerate}
Then we can obtain the above improved symmetric monoidal category as a decategorification of the symmetric monoidal bicategory $H(\mathbb{F}\textnormal{Cospan}(\bold{C}))$.

>>>>

\begin{abstract}
\noindent
The first two authors have developed a compositional framework well-suited for studying networks that are built out of finite sets equipped with extra stuff. This framework, which goes by the name of `structured cospans', utilizes double categories where the objects are seen as inputs and outputs, morphisms are `open networks', and 2-morphisms are maps between open networks. In this setup a functor $L \colon \textrm{A} \to \textrm{X}$, which is typically a left adjoint, is used to replace the objects and vertical 1-morphisms of a given double category $\mathbb{X}$ with the objects and morphisms, respectively, of the category $\textrm{A}$. Horizontal 1-cells are then cospans in $\textrm{X}$ of a particular form with 2-morphisms given by maps of these cospans. Fong has also developed a similar framework which also uses cospans and goes by the name of `decorated cospans'. In this setup, a lax monoidal functor $F \colon \textrm{A} \to \textrm{Set}$ is used to `decorate' the apices of cospans in $\textrm{A}$ with elements of $\textrm{Set}$ giving the cospans extra structure. Using a slight variation of Fong's framework, we prove that these two frameworks are equivalent in the situation where a left adjoint can be obtained from a lax monoidal pseudofunctor using a well known construction of Grothendieck.
\end{abstract}

\section{Introduction}
Networks are playing an increasingly prominent role in our understanding of the world and as a result, methods of characterizing and studying networks have become evermore necessary. Applied category theory provides such an avenue and much work has been done towards this effort  \cite{BC,BCR,BF,BFP,BP,Cour,Fong,JM,LS,Yass}. One of the more recent frameworks developed suitable for studying networks is Fong's `decorated cospans' \cite{Fong}. A $\textbf{cospan}$ in any category is a diagram of the form:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,1) {$c$};
\node (C) at (2,0) {$b$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
We call $c$ the $\textbf{apex}$ of the cospan and $a$ and $b$ the $\textbf{feet}$ of the cospan. The morphisms $i$ and $o$ are called the $\textbf{legs}$. Cospans are ideal for realizing networks and more generally `open' systems where here open means each system or network comes with prescribed inputs and outputs which are represented by the feet of the cospan $a$ and $b$, respectively. Two open systems viewed as cospans such that the outputs of the first and the inputs of the second coincide can then naturally be composed via pushout. For example, we can compose the above cospan with the following:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$b$};
\node (B) at (1,1) {$f$};
\node (C) at (2,0) {$e$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i^\prime$} (B)
(C) edge node[above]{$o^\prime$} (B);
\end{tikzpicture}
\]
to obtain a new cospan whose left foot is the left foot of the first, whose right foot is the right foot of the second, and whose apex and legs are given by the pushout.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,1) {$c$};
\node (C) at (2,0) {$b$};
\node (D) at (3,1) {$f$};
\node (E) at (4,0) {$e$};
\node (F) at (2,2) {$c+_b f$};
\node (G) at (2,1.25) {$c+f$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(C) edge node[above]{$i^\prime$} (D)
(E) edge node[above]{$o^\prime$} (D)
(B) edge node[left]{$\psi j_c$} (F)
(D) edge node[right]{$\psi j_f$} (F)
(B) edge node[above]{$j_c$} (G)
(D) edge node[above]{$j_f$} (G)
(G) edge node [left] {$\psi$} (F);
\end{tikzpicture}
\]

In Fong's theory of decorated cospans, given a category $\textrm{A}$ with finite colmits, a symmetric lax monoidal functor $F \colon (\textrm{A},+,0) \to (\textrm{Set},\times,1)$ is used to `decorate' the apex of a cospan in $\textrm{A}$ with an element of the image of its apex under the functor $F$, in the case of the first cospan above, a morphism $d \colon 1 \to F(c)$. Here $F(c)$ is the collection of all $F$-decorations on the object $c$ and the morphism $d$ is selecting a particular one. Thus a decorated cospan is given by a pair:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c$};
\node (C) at (2,0) {$b$};
\node (D) at (3,0) {$I$};
\node (E) at (4,0) {$F(c)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(D) edge node [above] {$d$} (E);
\end{tikzpicture}
\]
To obtain a decoration on the composition of two composable $F$-decorated cospans:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c$};
\node (C) at (2,0) {$b$};
\node (F) at (3,0) {$b$};
\node (G) at (4,0) {$f$};
\node (H) at (5,0) {$e$};
\node (D) at (0.5,-0.5) {$I$};
\node (E) at (1.5,-0.5) {$F(c)$};
\node (D') at (3.5,-0.5) {$I$};
\node (E') at (4.5,-0.5) {$F(f)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(F) edge node[above]{$i^\prime$} (G)
(H) edge node[above]{$o^\prime$} (G)
(D) edge node [above] {$d_1$} (E)
(D') edge node [above] {$d_2$} (E');
\end{tikzpicture}
\]
we use the natural map from a coproduct to a pushout as well as the natural map that comes as part of the structure of a lax monoidal functor:
 $$1 \xrightarrow{\lambda^{-1}} 1 \times 1 \xrightarrow{d_1 \times d_2} F(c) \times F(f) \xrightarrow{\phi_{c,f}} F(c+ f) \xrightarrow{F(j_{c,f})} F(c+_b f).$$

From this symmetric lax monoidal functor $F \colon \textrm{A} \to \mathrm{Set}$, Fong then constructs a symmetric monoidal category $F\textnormal{Cospan}(\mathrm{A})$ which has:
\begin{enumerate}
\item{objects as those of $\mathrm{A}$ and}
\item{morphisms as isomorphism classes of $F$-decorated cospans, where an $F$-decorated cospan is given as above, and two $F$-decorated cospans are in the same isomorphism class if there exists an isomorphism between the apices such that the following diagrams commute:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,1) {$c$};
\node (B') at (1,-1) {$c^\prime$};
\node (C) at (2,0) {$b$};
\node (D) at (3,0) {$1$};
\node (E) at (4,0.5) {$F(c)$};
\node (E') at (4,-0.5) {$F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[below]{$i^\prime$} (B')
(C) edge node[below]{$o^\prime$} (B')
(B) edge node [left] {$f$} (B')
(D) edge node [above] {$d$} (E)
(D) edge node [below] {$d^\prime$} (E')
(E) edge node [right] {$F(f)$} (E');
\end{tikzpicture}
\]
}
\end{enumerate}
There are some subtleties to this approach. To illustrate, consider the example where $F \colon \textrm{FinSet} \to \textrm{Set}$ is the symmetric lax monoidal functor that assigns to a finite set $N$ the (large) set of all possible graph structures on the finite set $N$, where a graph structure on $N$ is given by a diagram in $\textrm{Set}$ of the form:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$E$};
\node (B) at (1,0) {$N.$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge[bend left] node[above]{$s$} (B)
(A) edge[bend right] node[below]{$t$} (B);
\end{tikzpicture}
\]
Let $N=\{ v_1,v_2 \}$ be a two element set. Then one element of the set $F(N)$, in other words, one possible decoration, which is a graph structure, on the set $N$ is given by a single edge $e$ whose source and target are $v_1$ and $v_2$, respectively.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$v_1$};
\node (B) at (1,0) {$v_2$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$e$} (B);
\end{tikzpicture}
\]
Denote this element of $F(N)$ as $d$. We could label the edge $e$ something else, for instance, $e^\prime$, and this single edge would another element of $F(N)$.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$v_1$};
\node (B) at (1,0) {$v_2$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$e^\prime$} (B);
\end{tikzpicture}
\]
Denote this second element of $F(N)$ as $d^\prime$. These two graphs constitute distinct isomorphism classes in the symmetric monoidal category $F$Cospan$(\textrm{FinSet})$, as $f=\id_N$ and thus $F(f)=\id_{F(N)}$, but clearly $d \neq d^\prime$. The issue here is that the functor $F$ is decorating each finite set $N$ not only with extra structure, but with extra \emph{stuff}. Moreover, the finite set $N$ is being decorated with \emph{different} stuff in each instance and this difference in stuff is not entirely detected by the function $F(f)$. This phenomenon doesn't occur when the decorations only involve extra structure, and this has been utilized in other works \cite{BFP,BP,Yass}.

One approach to remedying this nuisance is to instead view $F$ as a functor $F \colon \textrm{A} \to \bold{Cat}$ and then consider $F(c)$ not as a \emph{set} of decorations on the element $c$ but as a \emph{category} of decorations on the element $c$. Then one can construct a similar symmetric monoidal category similar to Fong's, which will be called $D(H(\mathbb{F}\textnormal{Cospan}(\mathrm{A})))$ for reasons that will be explained later. This symmetric monoidal category will have:
\begin{enumerate}
\item{objects as those of $\mathrm{A}$ and}
\item{morphisms as isomorphism classes of cospans of $\mathrm{A}$ together with an element of image of the apex under the functor $F$.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c$};
\node (C) at (2,0) {$b$};
\node (D) at (3,0) {$d \in F(c)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
Except now, given another morphism:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c^\prime$};
\node (C) at (2,0) {$b$};
\node (D) at (3,0) {$d^\prime \in F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i^\prime$} (B)
(C) edge node[above]{$o^\prime$} (B);
\end{tikzpicture}
\]
these two morphisms are in the same isomorphism class if the following diagrams commute
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,1) {$c$};
\node (B') at (1,-1) {$c^\prime$};
\node (C) at (2,0) {$b$};
\node (D) at (3,0) {$1$};
\node (E) at (4,0.5) {$F(c)$};
\node (E') at (4,-0.5) {$F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[below]{$i^\prime$} (B')
(C) edge node[below]{$o^\prime$} (B')
(B) edge node [left] {$f$} (B')
(D) edge node [above] {$d$} (E)
(D) edge node [below] {$d^\prime$} (E')
(E) edge node [right] {$F(f)$} (E');
\end{tikzpicture}
\]
and there exists an isomorphism $\iota \colon F(f)(d) \to d^\prime$ in the category $F(c^\prime)$.
}
\end{enumerate}
Note that the existence of an isomorphism $\iota \colon F(f)(d) \to d^\prime$ is equivalent to the triangle above on the right commuting up to a 2-isomorphism.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (4,0) {$\iota \Swarrow$};
\node (D) at (3,0) {$1$};
\node (E) at (4.5,1) {$F(c)$};
\node (E') at (4.5,-1) {$F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(D) edge node [above] {$d$} (E)
(D) edge node [below] {$d^\prime$} (E')
(E) edge node [right] {$F(f)$} (E');
\end{tikzpicture}
\]
In other words, the isomorphism $\iota$ is a \emph{natural} isomorphism. We will typically just write this as $\iota \colon F(f)(d) \to d^\prime$ to conserve space.

In this new category, the added structure of an isomorphism in the category of decorations shapes the isomorphism classes to look as one would expect. Fong's symmetric monoidal category has been extended to a bicategory by the first author \cite{Cour} and this was one of the original illuminations of the above obstacle. In this bicategory, objects and morphisms are given as in Fong's category modulo taking cospans up to isomorphism class and 2-morphisms are given by a pair of commuting diagrams:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,1) {$c$};
\node (B') at (1,-1) {$c^\prime$};
\node (C) at (2,0) {$b$};
\node (D) at (3,0) {$1$};
\node (E) at (4,0.5) {$F(c)$};
\node (E') at (4,-0.5) {$F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[below]{$i^\prime$} (B')
(C) edge node[below]{$o^\prime$} (B')
(B) edge node [left] {$f$} (B')
(D) edge node [above] {$d$} (E)
(D) edge node [below] {$d^\prime$} (E')
(E) edge node [right] {$F(f)$} (E');
\end{tikzpicture}
\]
In this case, there would be no 2-morphism from:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$v_1$};
\node (B) at (1,0) {$v_2$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$e$} (B);
\end{tikzpicture}
\]
to:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$v_1$};
\node (B) at (1,0) {$v_2.$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$e^\prime$} (B);
\end{tikzpicture}
\]
Morally speaking, these two simple single-edged graphs should belong to the same isomorphism class, as there is a natural isomorphism of graphs between them.

In the present work, we construct a symmetric monoidal `double category' $\mathbb{F}\textnormal{Cospan}(\textrm{A})$. Double categories were first introduced by Ehresmann \cite{Ehresmann63, Ehresmann65} and symmetric monoidal double categories by Shulman \cite{Shul}. They have long been used in topology and other branches of pure mathematics \cite{Brown1,Brown2}.  More recently they have been used to study open dynamical systems \cite{LS} and open Markov processes \cite{BC}. While a mere category has only objects and morphisms, a double category has a few more types of entities:

@@

We call $A, B, C$ and $D$ `objects', $f$ and $g$ `vertical 1-morphisms', $M$ and $N$ `horizontal 1-cells', and $a$ a `2-morphism'.   We can compose vertical 1-morphisms to get new vertical 1-morphisms and compose horizontal 1-cells to get new horizontal 1-cells.  We can compose the 2-morphisms in two ways: horizontally by setting squares side by side, and vertically by setting one on top of the other.   In a `strict' double category all these forms of composition are associative.  In a `pseudo' double category, horizontal 1-cells compose in a weakly associative manner: that is, the associative law holds only up to an invertible 2-morphism, called the `associator', which obeys a coherence law. The double category $\mathbb{F}\textnormal{Cospan}(\textrm{A})$ that we construct in this paper has:

@@

\item{objects as those of $\textrm{A}$,}
\item{vertical 1-morphisms as morphisms of $\textrm{A}$,}
\item{horizontal 1-cells as $F$-decorated cospans of $\textrm{A}$, which are pairs:

@@

We can then obtain a symmetric monoidal bicategory $H(\mathbb{F}\textnormal{Cospan}(\textrm{A}))$ as the `horizontal bicategory' of the symmetric monoidal double category $\mathbb{F}\textnormal{Cospan}(\textrm{A})$ using a result of Shulman \cite{Shul}. This bicategory will have:

@@

\item{objects as those of $\textrm{A}$,}

@@

\node (D) at (3,0) {$1$};

@@

\end{enumerate}
Then we can obtain the above improved symmetric monoidal category as a decategorification of the symmetric monoidal bicategory $H(\mathbb{F}\textnormal{Cospan}(\textrm{A}))$.

The outline for the paper is as follows: In the second section we present the definition of double category and symmetric monoidal double category, both of which can be found in earlier works \cite{Shul}. In the third section, we construct the symmetric monoidal double category $F\textnormal{Cospan}(\textrm{A})$ mentioned above as well as a result of Shulman that allows us to obtain a symmetric monoidal bicategory whose decategorification is an improved version of Fong's symmetric monoidal category. In the fourth section we investigate when a symmetric lax monoidal pseudofunctor gives rise to a left adjoint via the Grothendieck construction. In the fifth section we review another compositional framework which uses cospans equipped with extra structure due to the first two authors, namely `structured cospans' and prove that the double categorical versions of decorated cospans and structured cospans are equivalent, and in the final section, we present a few examples that can be realized with either framework and exhibit the equivalences.
